\pkg{fitdistrplus}: An \proglang{R} Package for Fitting Distributions

### 2014-02-05 JSS/fitdistrplus/paper2JSS.Rnw
new version
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@263 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

The \code{fitdistr} function estimates distribution parameters by maximizing the likelihood
function using the \code{optim} function. 
In some cases, other estimation methods could be prefered, 
such as maximum goodness-of-fit estimation (also called minimum distance estimation), as
proposed in the \proglang{R} package \pkg{actuar} with three different goodness-of-fit distances
\citep{actuarJSS}. 
While developping  the \pkg{fitdistrplus} package, a second objective  was to consider 
various estimation methods in addition to maximum likelihood estimation (MLE). Functions
were developped to enable moment matching estimation (MME), quantile matching estimation (QME), and 
maximum goodness-of-fit estimation (MGE) using eight different distances.
Moreover, the \pkg{fitdistrplus} package offers the possibility to specify a user-supplied function
for optimization, 
useful in cases where classical optimization techniques, not included in \code{optim}, are more adequate.

@@

Few packages on \sigle{CRAN} provide estimation procedures for any user-supplied parametric distribution 
and support different types of data. 
The \pkg{distrMod} package \citep{distrModJSS} provides an object-oriented (S4) 
implementation of probability models for non-censored data
and includes distribution fitting procedures for a given minimization criterion.
This package is based on the \proglang{R}-shipped function \code{mle} of the \pkg{stats4} package
producing S4 objects of class \code{"mle"} for which generic methods are 
implemented, e.g., \code{confint} and \code{logLik}.
In \pkg{fitdistrplus}, we chose to use the standard S3 class system for its understanding by most \proglang{R} users.
When designing the \pkg{fitdistrplus} package, we did not forget to implement
generic functions also available for S3 classes.
Finally, various other packages provide functions to estimate the mode, the moments or the L-moments of a distribution, see the reference manuals of  \pkg{modeest}, \pkg{lmomco} and \pkg{Lmoments} packages.

@@

%%% R code
<<quantilefitdist, echo=TRUE, fig=FALSE>>=
quantile(fendo.B, probs = 0.05)
quantile(ATV, probs = 0.05)
@

@@

A classical data set from the Danish insurance industry 
published in \cite{mcneil97} will be used to illustrate this method.
In \pkg{fitdistrplus}, the data set is stored in \code{danishuni} for the univariate version and contains
the loss amounts collected at Copenhagen Reinsurance between 1980 and 1990.
In actuarial science, it is standard to consider positive heavy-tailed distributions and have 
a special focus on the right-tail of the distributions.
In this numerical experiment,  the lognormal distribution is first considered and then the Pareto type II 
distribution \citep{Klugmanetal09}.

@@

As shown on Figure~\ref{fig:danish:mme}, MME and MLE fits
are far less distant (when looking at the right-tail) for the Pareto distribution 
than for the lognormal distribution on this data set. 
Furthermore, for these two distributions, the MME method better fits the
right-tail of the distribution, which seems logical since empirical moments 
are influenced by large observed values.
In the previous traces, we gave the values of goodness-of-fit statistics.
Whatever the statistic considered, the MLE-fitted lognormal always provides the best fit 
to the observed data.

@@

Quantile matching estimation (QME) is performed by setting the argument \code{method}
to \code{"qme"} in the call to \code{fitdist} and adding an argument \code{probs}
defining the probabilities for which the quantile matching is performed.
The length of this vector must be equal
to the number of parameters to estimate (as the vector of moment orders for MME).
Empirical quantiles are computed using the
\code{quantile} function of the \pkg{stats} package using  \code{type=7} 
by default (see \code{?quantile} and \cite{hyndmanfan96}).
But the type of quantile can be easily changed by using the \code{qty} 
argument in the call to the \code{qme} function.   
The quantile matching is carried out numerically, 
by minimizing the sum of squared differences between observed and theoretical quantiles.
%%% R code
<<danish.qme.echo, echo=TRUE, fig=FALSE>>=
fdanish.ln.QME1 <- fitdist(danishuni$Loss, "lnorm", method="qme", 
probs=c(1/3, 2/3))
fdanish.ln.QME2 <- fitdist(danishuni$Loss, "lnorm", method="qme", 
probs=c(8/10, 9/10))
cdfcomp(list(fdanish.ln.MLE, fdanish.ln.QME1, fdanish.ln.QME2), 
legend=c("MLE", "QME(1/3, 2/3)", "QME(8/10, 9/10)"), 
main="Fitting a lognormal distribution", xlogscale=TRUE, datapch=20)
@
Above is an example of fitting of a lognormal distribution to \code{danishuni} data set 
by matching probabilities $(p_1= 1/3, p_2=2/3)$ and $(p_1= 8/10, p_2=9/10)$.
As expected, the second QME fit gives more weight to the right-tail of the distribution.
%, despite we do not choose the Pareto type-II distribution.
Compared to the maximum likelihood estimation, the second QME fit best suits the right-tail
of the distribution, whereas the first QME fit best models the body of the distribution.
The quantile matching estimation is of particular interest when we need a focus 
around particular quantiles, e.g., $p=99.5\%$ in the Solvency II insurance context or
$p=5\%$ for the HC5 estimation in the ecotoxicology context.

@@

@@

The \proglang{R}  package \pkg{fitdistrplus} allows to easily fit distributions. 
Our main objective while developing this package was to provide tools for helping 
\proglang{R} users to fit distributions to data. 
We have been encouraged to pursue our work by feedbacks 
from users of our package in various areas as food or environmental risk assessment, epidemiology, ecology, 
molecular biology, genomics, bioinformatics, hydraulics, mechanics, financial and actuarial mathematics 
or operations research. Indeed, this package is already used by a lot of practionners and academics 
for simple MLE fits \citep{distrModJSS,jaloustreetal11,saketal11,kochetal12,marquetouxetal12,nordan12,
scholletal12,suuronenetal12,varoetal12,mandletal13}, for MLE fits and goodness-of-fit statistics 
\citep{garcia12,tarnczi11,bagariaetal12,benavidesetal12,breitbach12,poduje12,Pouillot10}, 
for MLE fits 
and bootstrap \citep{croucheretal12,meheustetal12,orellanoetal12,telloetal12,hoelzeretal12}, 
for MLE fits, 
bootstrap and goodness-of-fit statistics \citep{rebuge12}, 
for MME fit \citep{luangkesornetal12}, 
for 
censored MLE and bootstrap \citep{lehaetal11,poulliotetal12,jongenburgeretal12,commeauetal12,pouillotetal11}, 
for graphic analysing in \citep{anandetal12,rosa12}, 
for grouped-data fitting methods \citep{fusteinercostafreda12} 
or more generally \citep{eling12,busschaertetal10,brooksetal11,aktassjostrand11,eikhermann12,bakos11}.
Many extensions of this package are planned in the future: 
we target to extend to censored data some methods for the moment only available for non-censored data, 
especially concerning goodness-of-fit evaluation and fitting methods. We will also enlarge the choice of 
fitting methods for non-censored data, by proposing new goodness-of-fit distances (e.g., distances 
based on quantiles) for maximum goodness-of-fit estimation 
and new types of moments (e.g., limited expected values) for moment matching estimation. At last, we will
consider the case of multivariate distribution fitting.

\section{Acknowledgments}
\label{merci}
%%%%%%%%%%
The package would not have been at this stage without the stimulating contribution of R\'egis Pouillot  and 
Jean-Baptiste Denis, especially for its conceptualization. We also want to thank R\'egis Pouillot for 
his very valuable comments on the first version of this paper.

>>>>

The \code{fitdistr} function estimates distribution parameters by maximizing the likelihood
function using the \code{optim} function. 
No distinction between parameter role (e.g. a scale %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
parameter) is made, as our talk focus 
on parameter estimation from a general point-of-view. %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In some cases, other estimation methods could be prefered, 
such as maximum goodness-of-fit estimation (also called minimum distance estimation), as
proposed in the \proglang{R} package \pkg{actuar} with three different goodness-of-fit distances
\citep{actuarJSS}. 
While developping  the \pkg{fitdistrplus} package, a second objective  was to consider 
various estimation methods in addition to maximum likelihood estimation (MLE). Functions
were developped to enable moment matching estimation (MME), quantile matching estimation (QME), and 
maximum goodness-of-fit estimation (MGE) using eight different distances.
Moreover, the \pkg{fitdistrplus} package offers the possibility to specify a user-supplied function
for optimization, 
useful in cases where classical optimization techniques, not included in \code{optim}, are more adequate.

@@

Few packages on \sigle{CRAN} provide estimation procedures for any user-supplied parametric distribution 
and support different types of data. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The \pkg{distrMod} package \citep{distrModJSS} provides an object-oriented (S4)  
implementation of probability models and includes distribution fitting procedures  
for a given minimization criterion. 
This criterion is a user-supplied function 
which is sufficiently flexible to handle censored data, yet not in a trivial way,
see Example M4 of the \pkg{distrMod} vignette.
The fitting functions \code{MLEstimator} and \code{MDEstimator} return an S4 class 
for which a coercion method to class mle is provided so that the respective 
functionalities (e.g. \code{confint} and \code{logLik}) from package \pkg{stats4} 
are available, too.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In \pkg{fitdistrplus}, we chose to use the standard S3 class system for its understanding by most \proglang{R} users.
When designing the \pkg{fitdistrplus} package, we did not forget to implement
generic functions also available for S3 classes.
Finally, various other packages provide functions to estimate the mode, the moments or the L-moments of a distribution, see the reference manuals of  \pkg{modeest}, \pkg{lmomco} and \pkg{Lmoments} packages.

@@

%%% R code
<<quantilefitdist, echo=TRUE, fig=FALSE>>=
quantile(fendo.B, probs = 0.05)
quantile(ATV, probs = 0.05)
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In addition to the ecotoxicology context, the \code{quantile} generic function is also attractive in
the actuarial--financial context. In fact, the value-at-risk $VAR_\alpha$ is defined as the $1-\alpha$-quantile
of the loss distribution and can be computed with \code{quantile} on a \code{"fitdist"} object.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

A classical data set from the Danish insurance industry 
published in \cite{mcneil97} will be used to illustrate this method.
In \pkg{fitdistrplus}, the data set is stored in \code{danishuni} for the univariate version and contains
the loss amounts collected at Copenhagen Reinsurance between 1980 and 1990.
In actuarial science, it is standard to consider positive heavy-tailed distributions and have 
a special focus on the right-tail of the distributions.
In this numerical experiment, we choose classic actuarial 
distributions for loss modelling: the lognormal distribution and the Pareto type II 
distribution, see e.g. \citep{Klugmanetal09}.

@@

As shown on Figure~\ref{fig:danish:mme}, MME and MLE fits
are far less distant (when looking at the right-tail) for the Pareto distribution 
than for the lognormal distribution on this data set. 
Furthermore, for these two distributions, the MME method better fits the 
right-tail of the distribution from a visual point of view. 
This seems logical since empirical moments 
are influenced by large observed values.
In the previous traces, we gave the values of goodness-of-fit statistics.
Whatever the statistic considered, the MLE-fitted lognormal always provides the best fit 
to the observed data.

@@

Quantile matching estimation (QME) is performed by setting the argument \code{method}
to \code{"qme"} in the call to \code{fitdist} and adding an argument \code{probs}
defining the probabilities for which the quantile matching is performed.
The length of this vector must be equal
to the number of parameters to estimate (as the vector of moment orders for MME).
Empirical quantiles are computed using the
\code{quantile} function of the \pkg{stats} package using  \code{type=7} 
by default (see \code{?quantile} and \cite{hyndmanfan96}).
But the type of quantile can be easily changed by using the \code{qty} 
argument in the call to the \code{qme} function.   
The quantile matching is carried out numerically, 
by minimizing the sum of squared differences between observed and theoretical quantiles.
%%% R code
<<danish.qme.echo, echo=TRUE, fig=FALSE>>=
fdanish.ln.QME1 <- fitdist(danishuni$Loss, "lnorm", method="qme", 
probs=c(1/3, 2/3))
fdanish.ln.QME2 <- fitdist(danishuni$Loss, "lnorm", method="qme", 
probs=c(8/10, 9/10))
cdfcomp(list(fdanish.ln.MLE, fdanish.ln.QME1, fdanish.ln.QME2), 
legend=c("MLE", "QME(1/3, 2/3)", "QME(8/10, 9/10)"), 
main="Fitting a lognormal distribution", xlogscale=TRUE, datapch=20)
@
Above is an example of fitting of a lognormal distribution to \code{danishuni} data set 
by matching probabilities $(p_1= 1/3, p_2=2/3)$ and $(p_1= 8/10, p_2=9/10)$.
As expected, the second QME fit gives more weight to the right-tail of the distribution.
%, despite we do not choose the Pareto type-II distribution.
Compared to the maximum likelihood estimation, the second QME fit best suits the right-tail
of the distribution, whereas the first QME fit best models the body of the distribution.
The quantile matching estimation is of particular interest when we need to focus 
around particular quantiles, e.g., $p=99.5\%$ in the Solvency II insurance context or
$p=5\%$ for the HC5 estimation in the ecotoxicology context.

@@

Even if no error is raised when computing the optimization, changing the algorithm 
is of particular interest to enforce bounds on some parameters. For instance, a 
volatility parameter $\sigma$ is strictly positive $\sigma>0$ and a probability 
 parameter $p$ lies in $p\in [0,1]$. This is possible by using arguments \code{lower} 
 and/or \code{upper}, for which their use automatically forces 
 \code{optim.method="L-BFGS-B"}.

@@

The \proglang{R}  package \pkg{fitdistrplus} allows to easily fit distributions. 
Our main objective while developing this package was to provide tools for helping 
\proglang{R} users to fit distributions to data. 
We have been encouraged to pursue our work by feedbacks 
from users of our package in various areas as food or environmental risk assessment, epidemiology, ecology, 
molecular biology, genomics, bioinformatics, hydraulics, mechanics, financial and actuarial mathematics 
or operations research. Indeed, this package is already used by a lot of practionners and academics 
for simple MLE fits \citep{jaloustreetal11,saketal11,kochetal12,marquetouxetal12,nordan12,
scholletal12,suuronenetal12,varoetal12,mandletal13,mala2013,mala2013publi,nadarajahbakar2013,simoetal13,
westphalfitch13,wayland13,vianaetal13,wu13,staggeetal13,fiorellietal13,mukhopadhyay13,tikoleetal13,
voigtetal14}, 
for MLE fits and goodness-of-fit statistics 
\citep{tarnczi11,garcia12,bagariaetal12,benavidesetal12,breitbach12,poduje12,Pouillot10,vaninsky13}, 
for MLE fits and bootstrap \citep{croucheretal12,meheustetal12,orellanoetal12,telloetal12,hoelzeretal12,
prosseretal13,Zhang2013,lebaut13,Rigaux2014}, 
for MLE fits, bootstrap and goodness-of-fit statistics \citep{rebuge12,larrasetal13}, 
for MME fit \citep{luangkesornetal12,callauetal13,satoetal13}, 
for censored MLE and bootstrap \citep{lehaetal11,pouillotetal11,poulliotetal12,jongenburgeretal12,
commeauetal12,Kingetal2013,contrerasetal2013,duarte13,feng13}, 
for graphic analysing in \citep{anandetal12,rosa12}, 
for grouped-data fitting methods \citep{fusteinercostafreda12} 
or more generally \citep{busschaertetal10,brooksetal11,aktassjostrand11,bakos11,eling12,eikhermann12,
sosaetal2013,srinivasanetal2013,meyeretal13,Guillier2013471,Daelmanetal13,gustafssonetal13,
eiketal13,Wu2:2013,drakeetal2014}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The \pkg{fitdistrplus} package is also complementary with the \pkg{distrMod} package, \citep{distrModJSS}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Many extensions of this package are planned in the future: 
we target to extend to censored data some methods for the moment only available for non-censored data, 
especially concerning goodness-of-fit evaluation and fitting methods. We will also enlarge the choice of 
fitting methods for non-censored data, by proposing new goodness-of-fit distances (e.g., distances 
based on quantiles) for maximum goodness-of-fit estimation 
and new types of moments (e.g., limited expected values) for moment matching estimation. At last, we will
consider the case of multivariate distribution fitting.

\section{Acknowledgments}
\label{merci}
%%%%%%%%%%
The package would not have been at this stage without the stimulating contribution of R\'egis Pouillot  and 
Jean-Baptiste Denis, especially for its conceptualization. We also want to thank R\'egis Pouillot for 
his very valuable comments on the first version of this paper.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The authors gratefully acknowledges the two anonymous referees and the Editor for useful and constructive
comments. The remaining errors, of course, should be attributed to the authors alone.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
